---
title: "Environmental data prep"
author: Nick McManus
output: html_document
date: '2023-09-28'
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)    ## always
library(here)         ## reading/writing data
library(purrr)        ## run iterative fxns faster
library(sf)           ## vector objects
library(terra)        ## Better/faster GIS
library(raster)       ## GIS format required by Dismo
library(tigris)       ## CA boundary shpfile
library(FedData)      ## Downloading SSURGO & NLCD data
```


This script is used for generating and prepping environmental data to be used in the SDM script.


# BCMv8 Data

### Convert ASC to TIF

Publicly available BCMv8 files are in .asc format. ASCII files have associated coordinates but usually do not have an explicitly defined CRS when read in as a raster. Because the SDM will be run in a samples with data (SWD) format, .asc files are not necessary to run MaxEnt. Additionally, .tif have smaller file sizes than .asc, so conversion saves on data storage. The `asc_to_tif()` function converts an entire directory of .asc files to .tif files with appropriate CR. BCM data uses NAD83 CA Albers (EPSG:3310), so this is the default for the function.

BCMv8 data can be directly downloaded from USGS at:
https://www.sciencebase.gov/catalog/item/5f29c62d82cef313ed9edb39

```{r}
## Read in fxn
source(here('R/asc_to_tif.R'))

## Assign file path (selects all .asc files in directory)
filepath = here('data/bcmv8/2000_2022//')

## Run fxn
asc_to_tif(filepath, remove = TRUE)
```



### Quarterly rasters

Using the `quarter_rast()` function, we'll generate rasters of either mean or cumulative values across the summer (Jun-Aug) or winter (Dec-Feb) quarter for each water year. These will then be used as additional environmental variables for extraction. 
```{r}
## Read in fxn
source(here("R/quarterly_rast.R"))

## Define fxn variables
pathIn = here("data/bcmv8/2000_2022//")
pathOut = here("data/bcmv8/quarterly_avgs//")
startYear = 2000
endYear = 2022

## Run for winter ppt
quarter_rast(var="ppt", quarter="winter", method="mean",
            startYear, endYear,
            pathIn, pathOut)

quarter_rast(var="ppt", quarter="winter", method="sum",
            startYear, endYear,
            pathIn, pathOut)

## Run for summer tmx
quarter_rast(var="tmx", quarter="summer", method = "mean",
            startYear, endYear,
            pathIn, pathOut)

```


# gNATSGO Data
Due to the large number of species occurrences within NA pockets of California gSSURGO data, we'll be using the gridded National Soil Survey Geographic Database (gNATSGO) for extracting soil data. NATSGO data combines SSURGO and STATSGO, prioritizing SSURGO data where available and filling in missing areas with lower-level STATSGO data. gNATSGO also differs from gSSURGO in that mapunit level data is spatialized as a 10m resolution raster rather than a vector. Due to the resolution of other environmental data and the coordinate precision with which species occurrence data is filtered, this raster should be satisfactory for extracting soil characteristics. 

gNATSGO data by state can be directly downloaded as an ESRI geodatabase (.gdb) from the USDA's Natural Resources Conservation Service (NRCS):
https://nrcs.app.box.com/v/soils/folder/191785692827

More information about the database can be found on the NRCS' website:
https://www.nrcs.usda.gov/resources/data-and-reports/gridded-national-soil-survey-geographic-database-gnatsgo#download


**NOTE:** While tabular data from .gdb can be read into R, I have not yet found a method for directly reading in the mapunit raster. Currently, this raster is obtained by opening the .gdb in ArcGIS Pro and exporting the raster as a .tif file.

The code and workflow for preparing gSSURGO data is still included in the following section for future reference. The workflow largely remains the same for both data sources.

Reading in data from gNATSGO database saved locally
```{r}
## Read in using sf package to specify file in database
## Comes in as df
horizon = st_read(here("data/natsgo/gNATSGO_CA/gNATSGO_CA.gdb"), 
                   layer = "chorizon")
component = st_read(here("data/natsgo/gNATSGO_CA/gNATSGO_CA.gdb"), 
                    layer = "component")
mapunit = st_read(here("data/natsgo/gNATSGO_CA/gNATSGO_CA.gdb"), 
                  layer = "mapunit")

## Read in fxn
source(here('R/natsgo.R'))

natsgo_avg(horizon, component, mapunit, 
           depth = 200, pathOut = here("data/natsgo//"))
```

Briefly exploring breakdown of component-level soil taxonomy 
```{r, eval=FALSE}
taxon_counts <- component %>% 
  mutate(compkind = as.factor(compkind)) %>% 
  filter(majcompflag == "Yes") %>% 
  group_by(compkind) %>% 
  summarize(n = length(compkind))

## How many components for each series?
series_counts <- component %>% 
  mutate(compname = as.factor(compname)) %>% 
  filter(compkind == "Series") %>% 
  filter(majcompflag == "Yes") %>% 
  group_by(compname) %>% 
  summarize(n = length(compname))

# ## Quick check on number of unique values to
# ## anticipate joining situation
# length(unique(component$cokey))
# [1] 93614
# length(unique(chorizon$cokey))
# [1] 43989
# ## Quite a few components w/o horizon level data
# length(unique(component$mukey))
# [1] 19258
# length(unique(mapunit$mukey))
# [1] 19258
# ## Matching number of mapunits (expected)
```


### Test code:
Trying `arcgisbinding` package to read in raster but not working. Maybe look into this more later, but for now just manually read in .gdb and exported raster from ArcGIS.
```{r}
## Install and load pkg
install.packages("arcgisbinding", repos="https://r.esri.com", type="win.binary")
library(arcgisbinding)
## Need to run first to check ESRI license on computer
arc.check_product()

raster <- arc.open(here("data/ssurgo/gNATSGO_CA/gNATSGO_CA/MapunitRaster_10m_CA_2023.tif"))

test <- arc.raster(arc.open(here("data/ssurgo/gNATSGO_CA/gNATSGO_CA.gdb")), "MapunitRaster_10m_CA_2023.tif")
```


# gSSURGO Data

Info and link to site go here.
https://websoilsurvey.nrcs.usda.gov/app/WebSoilSurvey.aspx

Directly downloading and extracting the most recent (2019) SSURGO data with the `FedData` package. 

Soil data from SSURGO is broken down into four different levels: soil survey areas, map units, components, and horizons. Data is downloaded by soil survey area, which is subdivided into map units. Spatial data (polygons) are provided only for map units. Each map unit contains one or more components, which helps break down map units by further soil type. Map unit and component data are related using a map unit key (mukey). Soil components are vertically broken up into horizons, representing a different strata. Horizon-level data contains the information we're interested for modeling species distribution, such as pH, percent organic matter, and cation exchange capacity. Horizon and component data are related using a component key (cokey). 

Because some components have many horizons, and some map units have several components, relating horizon-level data to map units for spatial extraction requires turning 1:many joins into 1:1 joins. This is done by selecting a cut-off soil depth, then finding a weighted average of horizon-level data based on the thickness of each horizon. Once this horizon data is joined to a component, we can use the percent area of a component w/in a map unit to find the weighted average again. 

**NOTE** `get_ssurgo()` supposedly can use spatial templates to download/crop/merge data for a given area, but this doesn't work with rasters or vectors of CA because the area is too big. I also tried reading in all soil areas at once, but it timed out before it could merge them all as one geopackage. The work-around is to download/extract each survey area in CA separately, then read in data from respective geopackages and merge. Not as nice as calling directly from a `get_ssurgo()` output (which I left commented out for future reference).

The following soil survey areas were omitted for lacking horizon-level data: CA663, CA704, CA793, CA804, CA806. 

### Download/merge data
First we'll download and extract all the soil data for CA locally as geopackages. Then we'll read in horizon, component, and map unit data for each soil area and merge them. 

Reading in data from SSURGO database:
```{r}
## List of soil survey areas in CA (w/horizon data)
CA_surveyAreas = c("AZ649", "AZ656", "CA011", "CA013", "CA021", "CA031", 
                   "CA033", "CA041", "CA053", "CA055", "CA067", "CA069", 
                   "CA077", "CA087", "CA095", "CA097", "CA101", "CA113", 
                   "CA600", "CA601", "CA602", "CA603", "CA604", "CA605", 
                   "CA606", "CA607", "CA608", "CA609", "CA610", "CA612",
                   "CA614", "CA618", "CA619", "CA620", "CA624", "CA628", 
                   "CA630", "CA632", "CA637", "CA638", "CA641", "CA642", 
                   "CA644", "CA645", "CA646", "CA647", "CA648", "CA649", 
                   "CA651", "CA653", "CA654", "CA659", "CA660", "CA664", 
                   "CA665", "CA666", "CA667", "CA668", "CA669", "CA670",
                   "CA671", "CA672", "CA673", "CA674", "CA675", "CA676",
                   "CA677", "CA678", "CA679", "CA680", "CA681", "CA682", 
                   "CA683", "CA684", "CA685", "CA686", "CA687", "CA688", 
                   "CA689", "CA691", "CA692", "CA693", "CA694", "CA695",
                   "CA696", "CA697", "CA698", "CA699", "CA701", "CA702", 
                   "CA703", "CA707", "CA708", "CA709", "CA713", "CA719", 
                   "CA724", "CA729", "CA731", "CA732", "CA740", "CA750",
                   "CA760", "CA763", "CA772", "CA776", "CA777", "CA788", 
                   "CA789", "CA790", "CA792", "CA794", "CA795", "CA796", 
                   "CA802", "CA803", "CA805")

## df w/get_ssurgo args for pmap
CA_df = data.frame(template = CA_surveyAreas,
                   label = CA_surveyAreas,
                   raw.dir = here("data/ssurgo/raw"),
                   extraction.dir = here("data/ssurgo/extracted"))

## Download/extract each survey area
## Only needs to be run once
## TAKES A LONG TIME! Commented out to be safe
# pmap(CA_df, get_ssurgo)


## Read in/merge horizon data
chorizon = data.frame()
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"), 
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "chorizon")
  chorizon = rbind(chorizon, x)
}

## Component data
component = data.frame()
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"), 
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "component")
  component = rbind(component, x)
}

## Mapunit data
mapunit = data.frame()
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"), 
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "mapunit")
  mapunit = rbind(mapunit, x)
}

```

```{r}
# ## **FOR REFERENCE**: If using fewer survey areas,
# ## use this code to automatically read in/merge dataset.
# ## Then filter/merge using same code below
# ssurgo = get_ssurgo(template = CA_surveyAreas, label = "CA")
# chorizon = ssurgo$tabular$chorizon
# component = ssurgo$tabular$component
# mapunit = ssurgo$tabular$mapunit
# ssurgo_sf = ssurgo$spatial
```


### Filter and join
Perform series of averages and joins to assign horizon data to each map unit.
```{r}
# ## Quick check on number of unique values to
# ## anticipate joining situation
# length(unique(component$cokey))
# [1] 91837
# length(unique(chorizon$cokey))
# [1] 42218
# ## Quite a few components w/o horizon level data
# length(unique(component$mukey))
# [1] 19126
# length(unique(mapunit$mukey))
# [1] 19126
# ## Matching number of mapunits (expected)


## Remove variables with all NAs
not_all_na <- function(x) any(!is.na(x)) #for data cleaning

chorizon <- chorizon %>% 
  select_if(not_all_na)
component <- component %>% 
  select_if(not_all_na)
mapunit <- mapunit %>% 
  select_if(not_all_na)


## Finding total soil depth.... not sure why this is a step
depth = chorizon %>% 
  group_by(cokey) %>% 
  summarize(total_depth = max(hzdepb.r))

## Remove horizons that start below 1m
chorizon200 = chorizon %>% 
  filter(hzdept.r < 200) %>% 
  droplevels()

## How many components still have more than 1 horizon
## after filtering in last step?
nrow(
  chorizon200 %>% 
    group_by(cokey) %>% 
    summarize(count = n()) %>% 
    filter(count > 1)
  )
# [1] 31000

## To deal w/this, we'll summarize var of interest
## w/weighted mean of horizon thickness (above 30cm cutoff)
chorizon200_wmean = chorizon200 %>% 
  ## find thickness of each horizon
  mutate(thick = ifelse(hzdepb.r > 200, 200 - hzdept.r,
                        hzdepb.r - hzdept.r)) %>% 
  ## weighted mean of each variable by component
  group_by(cokey) %>% 
  summarize(om = round(weighted.mean(om.r, thick, na.rm = TRUE), 2),
            cec = round(weighted.mean(cec7.r, thick, na.rm = TRUE), 2),
            ph = round(weighted.mean(ph1to1h2o.r, thick),2)) %>% 
  ## join w/depth
  left_join(., depth, by = "cokey")


## Filter component data for variables of interest
component = component %>% 
  dplyr::select(c(comppct.r, compname, slope.r, mukey, cokey))

## join with horizon data
component_horizon = left_join(component, chorizon200_wmean, by = "cokey")

## Find weighted average of variables based on % component in a mapunit
full_soil = component_horizon %>% 
  group_by(mukey) %>% 
  summarize(om = round(weighted.mean(om, comppct.r, na.rm = TRUE),2),
            cec = round(weighted.mean(cec, comppct.r, na.rm = TRUE), 2),
            ph = round(weighted.mean(ph, comppct.r, na.rm = TRUE),2),
            slope = round(weighted.mean(slope.r, comppct.r, na.rm = TRUE), 2)) %>% 
  ## join w/mapunit data
  left_join(., mapunit, by = "mukey")

## remove mapunit variables we don't care about and
## convert commas to _ in muname so its csv compatible 
full_soil <- full_soil %>% 
  mutate(muname = gsub(", ", "_", muname)) %>% 
  dplyr::select(!c(mukind, mapunitlfw.l:mapunitlf.h)) %>% 
  ## convert mukey from dbl to char for spatial join step
  mutate(mukey = as.character(mukey))

## Save as intermediate CSV
write_csv(full_soil, here("data/ssurgo/horizon_CA.csv"))
```


### Spatialize and export
```{r}
## Merge spatial and map unit data 
## first read in/merge all spatial data (will take a while!)
ssurgo_sf = NULL
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"),
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "geometry")
  ssurgo_sf = rbind(ssurgo_sf, x)
}

## Merge with our final soil data
ssurgo_sf <- ssurgo_sf %>% 
  janitor::clean_names() %>% 
  left_join(., full_soil, by = c("mukey", "musym"))

## Create separate sf for variables to be extracted
ssurgo_sf_filter <- ssurgo_sf %>% 
  dplyr::select(om, ph, cec)

## Save as .shp
st_write(ssurgo_sf_filter, here("data/ssurgo/ssurgo_ca_select.shp"))

## convert to vect first
## some variable names changed by st_write...
ssurgo_vect <- vect(ssurgo_sf)
writeVector(ssurgo_vect, here("data/ssurgo/ssurgo_ca_full.shp"))
```


# NLCD
Directly download NLCD data using the `FedData` package and a shapefile of CA provided by the `tigris` package. 
```{r}
## Get CA boundary from 2021 census info
ca_shp <- tigris::states(resolution = "500k") %>% 
  janitor::clean_names() %>% 
  filter(name == "California")

## Download in NLCD data
years <- c("2001", "2004", "2006", "2008", 
           "2011", "2013", "2016", "2019", "2021")

for (i in 1:length(years)) {
  nlcd <- FedData::get_nlcd(template = ca_shp, label = "CA",
                            year = years[i],
                            dataset = "landcover",
                            extraction.dir = here("data/"))
}

```

Exploring how many sppOcc are on high/med intensity development?
```{r}
## List only files with .tif extension (e.g. not .tif.aux)
rast_list <- grep(list.files(here("data/nlcd/rasters")), 
              pattern = ".tif.", invert = TRUE, value = TRUE)
rast_yrs <- as.numeric(regmatches(rast_list, regexpr("[0-9].*[0-9]", rast_list)))
rast_list_df <- data.frame(rast=rast_list,
                           year=rast_yrs)

## Sample spp to test
names <- c("a_polycarpa", "a_menziesii", "p_ciliata", "l_pentachaeta")

## Loop through land-use extractions at occ point
## Output in tidy table
nlcd_extract <- data.frame()
plant_ext<- data.frame()
# p in 1:length(names)
for(p in 1:length(names)){
  x <- read_csv(paste0(here("data/background/occ/combined_spp_occ//"), 
                      names[p], "_lowFilter.csv"))
  plant_ext <- data.frame()
  
      for(i in 1:nrow(rast_list_df)){
        r <- rast(paste0(here("data/nlcd/rasters//"), rast_list_df$rast[i]))
        if(i != nrow(rast_list_df)) {
          x_range <- x %>% 
                 filter(year >= rast_list_df$year[i] 
                        & year < rast_list_df$year[i+1])
        ##If filtering to last date, no next row to look at
        ##Just do 2021-present
        } else {
          x_range <- x %>% 
                 filter(year >= rast_list_df$year[i])
        }
        
        ##If no entries in that time period, then skip to next 
        if(nrow(x_range) > 0) {
          v <- vect(x_range, geom = c("lon", "lat"), crs = "WGS84")
          ext <- extract(r, v, ID = FALSE) %>% 
            janitor::clean_names() %>% 
            cbind(x_range, .)
        } else {
          next
        }
        
        ##One df of all extracts for plant
        plant_ext <- rbind(plant_ext, ext)
        
      }##End inner loop
  
  ##Summarize data and add to one consistent df
  ##nlcd_extract main output
  plants <- plant_ext %>%
    group_by(class) %>%
    summarize(count = length(class)) %>%
    mutate(count_pct = round(count/sum(.$count)*100,2),
           species = names[p])

  nlcd_extract <- rbind(nlcd_extract, plants)
    
}##End outer loop

## Save df for future reference
write_csv(nlcd_extract, here("data/nlcd/nlcd_occ_extract.csv"))

## Look at land use classes of interest
nlcd_summary <- nlcd_extract %>% 
  filter(class %in% c("Developed, Low Intensity",
                      "Developed, Medium Intensity",
                      "Developed High Intensity",
                      "Cultivated Crops")) 

## Results in table for sharing
library(kableExtra)
nlcd_summary %>% 
  filter(!species %in% c("a_polycarpa")) %>% 
  kbl() %>% 
  kable_styling(full_width=F, bootstrap_options = "striped")
```


# FMMP
Farmland Mapping & Monitoring Program data is used to explore more detailed farmland classifications. This data is downloaded directly from the CA Dept of Conservation:
https://gis.conservation.ca.gov/portal/home/group.html?id=b1494c705cb34d01acf78f4927a75b8f#overview

Newer years (2016-present) format data as .gdb, while older years just provide all .shp files in one folder. The following code chunk is used to extract all the layers (counties) from the newer .gdb data, isolate prime farmland, and save as one .shp file for the year
```{r}
save_prime_fxn <- function(year) {
  ## Read in all layers (counties) of .gdb
  path <- paste0(here("data/fmmp//"), year, 
                 "/Important_Farmland_", year, ".gdb")
  lyrs <- st_layers(path)
  filter_prime <- function(lyr=lyrs$name) {
    x <- read_sf(dsn = path, layer=lyr) %>% 
      dplyr::filter(polygon_ty == "P")
  }
  prime_farms <- purrr::map(lyrs$name, filter_prime, .progress=TRUE)

  ## Combine all prime farmland from year into one sf
  counties <- data.frame()
  for (i in 1:length(prime_farms)){
    x <- prime_farms[[i]]
    counties <- rbind(counties, x)
  }
  
  ## Save .shp
  filename <- paste0(here("data/fmmp/prime_farmland/prime_farmland_"), 
                     year, ".shp")
  write_sf(counties, filename)
  
  ## Changes names when saving (cannot figure out why)
  ## So for making live easier, manually change all names
  ## and resave to match 2000-2014 data
  sf <- read_sf(paste0(here("data/fmmp/prime_farmland/prime_farmland_"), 
                     year, ".shp"))
  sf_rename <- sf %>% 
    rename(polygon_ac = plygn_c,
         county_nam = cnty_nm,
         upd_year = upd_yer,
         polygon_ty = plygn_t,
         Shape_Length = Shp_Lng,
         Shape_Area = Shap_Ar,
         Shape = geometry) %>% 
    vect()
  writeVector(sf_rename, (paste0(here("data/fmmp/prime_farmland/prime_farmland_"),
                                 year, ".shp")), overwrite=TRUE)
}

year <- c(2016, 2018, 2020)
purrr::map(years, save_prime_fxn, .progress=TRUE)
```

For 2000-2014, we'll do the same process, but the code slightly differs due to data format.
```{r}
## Overall fxn to read in, filter, merge, and write .shp for a year
prime_farm_filter <- function(year) {
  path <- paste0(here("data/fmmp//"), year, "/", year, "_FMMP_shape_files")
  lyrs <- list.files(path, pattern = ".shp", full.names = TRUE)
  fxn <- function(lyrs){
    x <- st_read(dsn = lyrs) %>% 
      dplyr::filter(polygon_ty=="P") %>% 
      janitor::clean_names()
  }
  prime_farms <- purrr::map(lyrs, fxn)
  
  ## Combine all prime farmland from year into one sf
  counties <- data.frame()
  for (i in 1:length(prime_farms)){
    x <- prime_farms[[i]]
    counties <- rbind(counties, x)
  }

  ## Save .shp
  filename <- paste0(here("data/fmmp/prime_farmland//"),
                     "prime_farmland_", year, ".shp")
  write_sf(counties, filename)
}##End fxn


## List of years
year = seq(2000,2014, by=2)
## Run fxn over all years
purrr::map(year, prime_farm_filter, .progress=TRUE)
```


How many spp occ are on prime farmland?
-**NOTE:** This loop doesn't fully work, because for a couple years (2016-2020) the extract returns multiple identical values (that aren't filtered out through `unique()` somehow). This returns an extraction df w/a different number of rows than the df of obs going into it, so they cannot be bound properly. 
```{r}
shp_list <- list.files(here("data/fmmp/prime_farmland/"), 
                          pattern = ".shp", full.names = TRUE )
year <- seq(2000, 2020, by=2)
names <- c("a_polycarpa", "a_menziesii", "p_ciliata", "l_pentachaeta")

## Loop through land-use extractions at occ point
## Output in tidy table
fmmp_extract <- data.frame()
# p in 1:length(names)
for(p in 1:length(names)){
  x <- read_csv(paste0(here("data/occ/combined_spp_occ//"), 
                      names[p], "_lowFilter.csv"))
  plant_ext <- data.frame()
    
    ## Extract for each year of FMMP data
    for(i in 1:length(shp_list)){
      shp <- vect(shp_list[i])
      
      ## For 2020 FMMP, just extract from 2020-2023
      if (i != length(shp_list)) {
        x_range <- x %>% 
               filter(year >= year[i] 
                      & year < year[i+1])
      } else {
        x_range <- x %>% 
               filter(year >= year[i])
      }
      
      ## Vectorize and extract obs w/in date range
      if(nrow(x_range) > 0) {
        v <- vect(x_range, geom = c("lon", "lat"), crs = "WGS84") %>% 
          project(y=crs(shp))
        ext <- extract(shp, v) %>% 
          janitor::clean_names() %>% 
          dplyr::select(polygon_ty) %>% 
          cbind(x_range, .)
      ## If no entries w/in date range, skip to next 
      } else {
        next
      }
      
      ##One df of all extracts for plant
      plant_ext <- rbind(plant_ext, ext)
      
    }##End inner loop

  ##Summarize data and add to one consistent df
  ##nlcd_extract main output
  plants <- plant_ext %>%
    mutate(prime = case_when(polygon_ty == "P" ~ 1),
           species = names[p]) %>% 
    group_by(species) %>%
    summarize(count = sum(prime, na.rm=TRUE),
              count_pct = round((sum(prime, na.rm=TRUE))/length(polygon_ty)*100,2))

  fmmp_extract <- rbind(fmmp_extract, plants)
    
}##End outer loop

write_csv(nlcd_extract, here("data/nlcd/nlcd_occ_extract.csv"))

## Look at some classes
nlcd_summary <- nlcd_extract %>% 
  filter(class %in% c("Developed, Low Intensity",
                      "Developed, Medium Intensity",
                      "Developed High Intensity",
                      "Cultivated Crops")) 

## Results in table for sharing
library(kableExtra)
nlcd_summary %>% 
  filter(!species %in% c("a_polycarpa")) %>% 
  kbl() %>% 
  kable_styling(full_width=F, bootstrap_options = "striped")
```
