---
title: "Environmental data prep"
author: Nick McManus
output: html_document
date: '2023-09-28'
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)    ## always
library(here)         ## reading/writing data
library(purrr)        ## run iterative fxns faster
library(sf)           ## vector objects
library(terra)        ## Better/faster GIS
library(raster)       ## GIS format required by Dismo
library(FedData)      ## Downloading SSURGO data
```


This script is used for generating and prepping environmental data to be used in the SDM. 

# Convert ASC to TIF

Publicly available BCMv8 files are in .asc format. ASCII files have associated coordinates but usually do not have an explicitly defined CRS when read in as a raster. Additionally, .tif have smaller file sizes than .asc, so conversion saves on data storage. The `asc_to_tif()` function converts an entire directory of .asc files to .tif files with appropriate CRS. BCM data uses NAD83 CA Albers (EPSG:3310), so this is the default for the function.
```{r}
## Read in fxn
source(here('R/asc_to_tif.R'))

## Assign file path (selects all .asc files in directory)
filepath = here('data/bcmv8/2000_2023//')

## Run fxn
asc_to_tif(filepath)
```



# BCM Quarterly rasters

Using the `quarter_rast()` function, we'll generate rasters of either mean or cumulative values across the summer (Jun-Aug) or winter (Dec-Feb) quarter for each water year. These will then be used as additional environmental variables for extraction. 
```{r}
## Read in fxn
source(here("R/quarterly_avg.R"))

## Define fxn variables
pathIn = here("data/bcmv8/2000_2022//")
pathOut = here("data/bcmv8/quarterly_avgs//")
startYear = 2000
endYear = 2022

## Run for winter ppt
quarter_rast(var="ppt", quarter="winter", method="mean",
            startYear, endYear,
            pathIn, pathOut)

quarter_rast(var="ppt", quarter="winter", method="sum",
            startYear, endYear,
            pathIn, pathOut)

## Run for summer tmx
quarter_rast(var="tmx", quarter="summer", method = "mean",
            startYear, endYear,
            pathIn, pathOut)

```



# SSURGO Data

Info and link to site go here.
https://websoilsurvey.nrcs.usda.gov/app/WebSoilSurvey.aspx


Directly downloading and extracting the most recent (2019) SSURGO data with the `FedData` package. 

Soil data from SSURGO is broken down into four different levels: soil survey areas, map units, components, and horizons. Data is downloaded by soil survey area, which is subdivided into map units. Spatial data (polygons) are provided only for map units. Each map unit contains one or more components, which helps break down map units by further soil type. Map unit and component data are related using a map unit key (mukey). Soil components are vertically broken up into horizons, representing a different strata. Horizon-level data contains the information we're interested for modeling species distribution, such as pH, percent organic matter, and cation exchange capacity. Horizon and component data are related using a component key (cokey). 

Because some components have many horizons, and some map units have several components, relating horizon-level data to map units for spatial extraction requires turning 1:many joins into 1:1 joins. This is done by selecting a cut-off soil depth, then finding a weighted average of horizon-level data based on the thickness of each horizon. Once this horizon data is joined to a component, we can use the percent area of a component w/in a map unit to find the weighted average again. 

**NOTE** `get_ssurgo()` supposedly can use spatial templates to download/crop/merge data for a given area, but this doesn't work with rasters or vectors of CA because the area is too big. I also tried reading in all soil areas at once, but it timed out before it could merge them all as one geopackage. The work-around is to download/extract each survey area in CA separately, then read in data from respective geopackages and merge. Not as nice as calling directly from a `get_ssurgo()` output (which I left commented out for future reference).

The following soil survey areas were omitted for lacking horizon-level data: CA663, CA704, CA793, CA804, CA806. 

### Download/merge data
First we'll download and extract all the soil data for CA locally as geopackages. Then we'll read in horizon, component, and map unit data for each soil area and merge them. 
```{r}
## List of soil survey areas in CA (w/horizon data)
CA_surveyAreas = c("AZ649", "AZ656", "CA011", "CA013", "CA021", "CA031", 
                   "CA033", "CA041", "CA053", "CA055", "CA067", "CA069", 
                   "CA077", "CA087", "CA095", "CA097", "CA101", "CA113", 
                   "CA600", "CA601", "CA602", "CA603", "CA604", "CA605", 
                   "CA606", "CA607", "CA608", "CA609", "CA610", "CA612",
                   "CA614", "CA618", "CA619", "CA620", "CA624", "CA628", 
                   "CA630", "CA632", "CA637", "CA638", "CA641", "CA642", 
                   "CA644", "CA645", "CA646", "CA647", "CA648", "CA649", 
                   "CA651", "CA653", "CA654", "CA659", "CA660", "CA664", 
                   "CA665", "CA666", "CA667", "CA668", "CA669", "CA670",
                   "CA671", "CA672", "CA673", "CA674", "CA675", "CA676",
                   "CA677", "CA678", "CA679", "CA680", "CA681", "CA682", 
                   "CA683", "CA684", "CA685", "CA686", "CA687", "CA688", 
                   "CA689", "CA691", "CA692", "CA693", "CA694", "CA695",
                   "CA696", "CA697", "CA698", "CA699", "CA701", "CA702", 
                   "CA703", "CA707", "CA708", "CA709", "CA713", "CA719", 
                   "CA724", "CA729", "CA731", "CA732", "CA740", "CA750",
                   "CA760", "CA763", "CA772", "CA776", "CA777", "CA788", 
                   "CA789", "CA790", "CA792", "CA794", "CA795", "CA796", 
                   "CA802", "CA803", "CA805")

## df w/get_ssurgo args for pmap
CA_df = data.frame(template = CA_surveyAreas,
                   label = CA_surveyAreas,
                   raw.dir = here("data/ssurgo/raw"),
                   extraction.dir = here("data/ssurgo/extracted"))

## Download/extract each survey area
pmap(CA_df, get_ssurgo)


## Read in/merge horizon data
chorizon = data.frame()
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"), 
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "chorizon")
  chorizon = rbind(chorizon, x)
}

## Component data
component = data.frame()
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"), 
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "component")
  component = rbind(component, x)
}

## Mapunit data
mapunit = data.frame()
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"), 
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "mapunit")
  mapunit = rbind(mapunit, x)
}

```

```{r}
# ## **FOR REFERENCE**: If using fewer survey areas,
# ## use this code to automatically read in/merge dataset.
# ## Then filter/merge using same code below
# ssurgo = get_ssurgo(template = CA_surveyAreas, label = "CA")
# chorizon = ssurgo$tabular$chorizon
# component = ssurgo$tabular$component
# mapunit = ssurgo$tabular$mapunit
# ssurgo_sf = ssurgo$spatial
```


### Filter and join
Perform series of averages and joins to assign horizon data to each map unit.
```{r}
# ## Quick check on number of unique values to
# ## anticipate joining situation
# length(unique(component$cokey))
# [1] 91837
# length(unique(chorizon$cokey))
# [1] 42218
# ## Quite a few components w/o horizon level data
# length(unique(component$mukey))
# [1] 19126
# length(unique(mapunit$mukey))
# [1] 19126
# ## Matching # of mapunits (expected)


## Remove variables with all NAs
not_all_na <- function(x) any(!is.na(x)) #for data cleaning

chorizon <- chorizon %>% 
  select_if(not_all_na)
component <- component %>% 
  select_if(not_all_na)
mapunit <- mapunit %>% 
  select_if(not_all_na)


## Finding total soil depth.... not sure why this is a step
depth = chorizon %>% 
  group_by(cokey) %>% 
  summarize(total_depth = max(hzdepb.r))

## Remove horizons that start below 30cm
chorizon30 = chorizon %>% 
  filter(hzdept.r < 30) %>% 
  droplevels()

## How many components still have more than 1 horizon
## after filtering in last step?
nrow(
  chorizon30 %>% 
    group_by(cokey) %>% 
    summarize(count = n()) %>% 
    filter(count > 1)
  )
# [1] 31000

## To deal w/this, we'll summarize var of interest
## w/weighted mean of horizon thickness (above 30cm cutoff)
chorizon30_wmean = chorizon30 %>% 
  ## find thickness of each horizon
  mutate(thick = ifelse(hzdepb.r > 30, 30 - hzdept.r,
                        hzdepb.r - hzdept.r)) %>% 
  ## weighted mean of each variable by component
  group_by(cokey) %>% 
  summarize(om = round(weighted.mean(om.r, thick, na.rm = TRUE), 2),
            cec = round(weighted.mean(cec7.r, thick, na.rm = TRUE), 2),
            ph = round(weighted.mean(ph1to1h2o.r, thick),2)) %>% 
  ## join w/depth
  left_join(., depth, by = "cokey")


## Filter component data for variables of interest
component = component %>% 
  dplyr::select(c(comppct.r, compname, slope.r, mukey, cokey))

## join with horizon data
component_horizon = left_join(component, chorizon30_wmean, by = "cokey")

## Find weighted average of variables based on % component in a mapunit
full_soil = component_horizon %>% 
  group_by(mukey) %>% 
  summarize(om = round(weighted.mean(om, comppct.r, na.rm = TRUE),2),
            cec = round(weighted.mean(cec, comppct.r, na.rm = TRUE), 2),
            ph = round(weighted.mean(ph, comppct.r, na.rm = TRUE),2),
            slope = round(weighted.mean(slope.r, comppct.r, na.rm = TRUE), 2)) %>% 
  ## join w/mapunit data
  left_join(., mapunit, by = "mukey")

## remove mapunit variables we don't care about and
## convert commas to _ in muname so its csv compatible 
full_soil <- full_soil %>% 
  mutate(muname = gsub(", ", "_", muname)) %>% 
  dplyr::select(!c(mukind, mapunitlfw.l:mapunitlfw.h)) %>% 
  ## convert mukey from dbl to char for spatial join step
  mutate(mukey = as.character(mukey))

## Save as intermediate CSV
write_csv(full_soil, here("data/ssurgo/horizon_CA.csv"))
```


### Spatialize and export
```{r}
## Merge spatial and map unit data 
## first read in/merge all spatial data (will take a while!)
ssurgo_sf = NULL
for (i in 1:nrow(CA_df)) {
  area = paste0(here("data/ssurgo/extracted//"),
                CA_df$template[i], 
                "_ssurgo.gpkg")
  
  x <- st_read(area, layer = "geometry")
  ssurgo_sf = rbind(ssurgo_sf, x)
}

## Merge with our final soil data
ssurgo_sf <- ssurgo_sf %>% 
  janitor::clean_names() %>% 
  left_join(., full_soil, by = c("mukey", "musym"))

## Create separate sf for variables to be extracted
ssurgo_sf_filter <- ssurgo_sf %>% 
  dplyr::select(om, ph, cec)

## Save as .shp
st_write(ssurgo_sf_filter, here("data/ssurgo/ssurgo_ca_select.shp"))

## convert to vect first
## some variable names changed by st_write...
ssurgo_vect <- vect(ssurgo_sf)
writeVector(ssurgo_vect, here("data/ssurgo/ssurgo_ca_full.shp"))
```


### Old code
Also found some rasters on Data Basin for pH (https://databasin.org/datasets/92c216fc8a0e41b9859b500de930ae7e/) and SOM (https://databasin.org/datasets/8cb1acdc738d4c9196435195b13bf972/).

Used this code to have rasters match the resolution and spatial extent of the the BCM layers, then export. 
As of now, not using because it's older (2008) and missing CEC. 
```{r}
## reference raster
bcm = rast(here("data/bcmv8/2000_2022/ppt2020jan.tif"))

## ssurgo rasters
ph = rast(here("data/ssurgo/ssurgo_ph.tif"))
som = rast(here("data/ssurgo/ssurgo_percent_som.tif"))

## change resolution from 90m to 270m (factor of 3) 
ph_270 <- aggregate(ph, fact=3, fun="mean", na.rm=TRUE) %>% 
  project(y = bcm)
## resample to ensure same extent
ph_resamp <- resample(ph_270, bcm, method="bilinear")
## export
writeRaster(ph_resamp, here("data/ssurgo/ssurgo_ph_270m.tif"))

## same with SOM
som_270 <- aggregate(som, fact=3, fun="mean", na.rm=TRUE) %>% 
  project(y=bcm)
som_resamp <- resample(som_270, bcm, method="bilinear")
writeRaster(som_resamp, here("data/ssurgo/ssurgo_som_270m.tif"))
```

