---
title: "Kern Natives SDM"
author: "Nick McManus"
date: "2023-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)    ## always
library(here)         ## reading/writing data
library(terra)        ## Better/faster GIS
library(raster)       ## GIS format required by Dismo
library(dismo)        ## Maxent pkg
library(lubridate)    ## Dates and progress bar
```


## Overview

This script extracts environmental data for each species occurrence (based on location and time). This data is then used to perform a species distribution model using Maxent with `dismo`.

*   Species occurrence data was pulled from GBIF (and maybe CalFlora?) and filtered to return reasonably accurate presence-only data in California. The script for this is in "src" directory.

*   Environmental data comes from the Basin Characterization Model version 8.0 (Flint et al., 2021) hosted by USGS. 



## Convert ASC to TIF

Publicly available BCMv8 files are in .asc format. ASCII files have associated coordinates but usually do not have an explicitly defined CRS when read in as a raster. Additionally, .tif have smaller file sizes than .asc, so conversion saves on data storage. Use this code to convert an entire file of .asc files to .tif files with appropriate CRS. BCM data uses NAD83 CA Albers (EPSG:3310), so this is the default for the function.
```{r}
## Read in fxn
source(here('R/asc_to_tif.R'))

## Assign file path (selects all .asc files in directory)
filepath = here('data/bcmv8/2021_2022/')

## Run fxn
asc_to_tif(filepath)
```




## Extract environmental data

Using BCMv8 rasters, we'll extract information for each species occurrence within the given timeframe. Outputs 

```{r}
## Read in fxn
source(here('R/env_extract.R'))
```

### Occurrence extraction:
```{r}
## Read in species occurrence df
sppOcc <- read_csv(here('data/species_occurrence_GBIF/a_polycarpa.csv')) %>% 
  janitor::clean_names() %>% 
  dplyr::select(gbifid, decimallatitude, decimallongitude, month, year)


### 2000-2009 Extraction ----------------------------------------
## select directory of env rasters
path <- here('data/bcmv8/2000_2009/')

## use fxn to extract env data at location of occurrences
occExtract_2000_2009 <- env_extract(2000, 2009, path, sppOcc)


### 2010-2020 Extraction ----------------------------------------
path <- here('data/bcmv8/2010_2020/')

occExtract_2010_2020 <- env_extract(2010, 2020, path, sppOcc)


### 2021-2022 Extraction ----------------------------------------
path <- here('data/bcmv8/2021_2022/')

occExtract_2021_2022 <- env_extract(2021, 2022, path, sppOcc)


### Merge and export --------------------------------------------
occExtract_2000_2022 <- rbind(occExtract_2000_2009, occExtract_2010_2020, occExtract_2021_2022)
write_csv(occExtract_2000_2022, here('data/swd/occExtract_2000_2022.csv'))
```


### Background extraction:
```{r}
## Read in background df
backOcc <- read_csv(here('data/species_occurrence_GBIF/background_points.csv')) %>% 
  janitor::clean_names()


### 2000-2009 Extraction ----------------------------------------
path <- here('data/bcmv8/2000_2009/')

backExtract_2000_2009 <- env_extract(2000, 2009, path, backOcc, lon = 'x', lat = 'y')


### 2010-2020 Extraction ----------------------------------------
path <- here('data/bcmv8/2010_2020/')

backExtract_2010_2020 <- env_extract(2010, 2020, path, backOcc, lon = 'x', lat = 'y')


### 2021-2022 Extraction ----------------------------------------
path <- here('data/bcmv8/2021_2022/')

backExtract_2021_2022 <- env_extract(2021, 2022, path, backOcc, lon = 'x', lat = 'y')


### Merge and export --------------------------------------------
backExtract_2000_2022 <- rbind(backExtract_2000_2009, backExtract_2010_2020, backExtract_2021_2022)
write_csv(backExtract_2000_2022, here('data/swd/backExtract_2000_2022.csv'))
```




## Maxent

```{r}

```


