---
title: "Species Distribution Model for Kern Natives"
author: "Nick McManus"
date: "2023-08-02"
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)    ## always
library(here)         ## reading/writing data
library(purrr)        ## run iterative fxns faster
library(sf)           ## vector objects
library(terra)        ## Better/faster GIS
library(raster)       ## GIS format required by Dismo
library(dismo)        ## Maxent pkg
library(rJava)        ## Needed for dismo
library(lubridate)    ## Dates and progress bar
library(corrplot)     ## Correlation matrix
library(FedData)      ## Downloading SSURGO data
```


# Overview

This script extracts environmental data for each species occurrence (based on location and time). This data is then used to perform a species distribution model using Maxent with `dismo`.

*   Species occurrence data was pulled from GBIF (and maybe CalFlora?) and filtered to return reasonably accurate presence-only data in California. The script for this is in "src" directory.

*   Environmental data comes from the Basin Characterization Model version 8.0 (Flint et al., 2021) hosted by USGS. 

*   SSURGO data info here....


# Extract environmental data

Using BCMv8 monthly and quarterly rasters, we'll extract information for each species occurrence (and background point) within the given time frame. The output will be one CSV with environmental data for each point.  
```{r}
## Read in fxn
source(here('R/env_extract.R'))

## Define fxn variables
pathMonth = here("data/bcmv8/2000_2022//")
pathQuarter = here("data/bcmv8/quarterly_avgs//")
soilRast = rast(here("data/natsgo/MapunitRaster_10m_CA_2023.tif"))
```


## Occurrence extraction:
```{r}
names <- c(
  "a_polycarpa",
  "p_arborea",
  "c_pungens",
  "l_pentachaeta",
  "p_ciliata",
  "a_menziesii",
  "c_lasiophyllus"
)

## Read in species occurrence df
##NOTE: starting with A.polycarpa as ref spp
sppOcc <- read_csv(paste0(here('data/occ/combined_spp_occ//'),
                          names[1],
                          "_lowFilter.csv"))

## Use fxn to extract env data at location of occurrences

### TESTING FOR SOIL DEPTHS
horizon30 = read_csv(here("data/natsgo/horizon30_CA.csv"))
occExtract_30 <- env_extract(startYear=2000, endYear=2022, 
                             pathMonth, pathQuarter, soilRast, 
                             horizon=horizon30, occ = sppOcc) %>% 
  dplyr::select(!c(mukey, musym:muname))
write_csv(occExtract_30, here('data/swd/a_polycarpa/a_polycarpa_30cm_lowFilter.csv'))

horizon100 = read_csv(here("data/natsgo/horizon100_CA.csv"))
occExtract_100 <- env_extract(startYear=2000, endYear=2022, 
                             pathMonth, pathQuarter, soilRast, 
                             horizon=horizon100, occ = sppOcc) %>% 
  dplyr::select(!c(mukey, musym:muname))
write_csv(occExtract_100, here('data/swd/a_polycarpa/a_polycarpa_100cm_lowFilter.csv'))

horizon200 = read_csv(here("data/natsgo/horizon200_CA.csv"))
occExtract_200 <- env_extract(startYear=2000, endYear=2022, 
                             pathMonth, pathQuarter, soilRast, 
                             horizon=horizon200, occ = sppOcc) %>% 
  dplyr::select(!c(mukey, musym:muname))
write_csv(occExtract_200, here('data/swd/a_polycarpa/a_polycarpa_200cm_lowFilter.csv'))

horizon.full = read_csv(here("data/natsgo/horizonFull_CA.csv"))
occExtract_full <- env_extract(startYear=2000, endYear=2022, 
                             pathMonth, pathQuarter, soilRast, 
                             horizon=horizon.full, occ = sppOcc) %>% 
  dplyr::select(!c(mukey, musym:muname))
write_csv(occExtract_full, here('data/swd/a_polycarpa/a_polycarpa_full_lowFilter.csv'))
```


## Background extraction:
First, we'll create a subset of the background points data that are located within 50km of the species occurrence points using the terra and sf packages. Then, we'll run the `env_extract()` fxn on this subset and export the results for use in a species-specific SDM.
```{r}
## Read in background df
backOcc <- read_csv(here('data/background/back_a_polycarpa_lowfilter_5km.csv')) %>% 
  janitor::clean_names()


## Use fxn to extract for background points
## NOTE: this will take a while to run
backExtract_30 <- env_extract(startYear=2000, endYear=2022,
                              pathMonth, pathQuarter, soilRast,
                              horizon=horizon30, occ = backOcc,
                              lon = "x", lat = "y") %>% 
  dplyr::select(!c(mukey, musym:muname))

backExtract_100 <- env_extract(startYear=2000, endYear=2022,
                               pathMonth, pathQuarter, soilRast,
                               horizon=horizon100, occ = backOcc,
                               lon = "x", lat = "y") %>% 
  dplyr::select(!c(mukey, musym:muname))

backExtract_200 <- env_extract(startYear=2000, endYear=2022,
                               pathMonth, pathQuarter, soilRast,
                               horizon=horizon200, occ = backOcc,
                               lon = "x", lat = "y") %>% 
  dplyr::select(!c(mukey, musym:muname))

backExtract_full <- env_extract(startYear=2000, endYear=2022,
                               pathMonth, pathQuarter, soilRast,
                               horizon=horizon.full, occ = backOcc,
                               lon = "x", lat = "y") %>% 
  dplyr::select(!c(mukey, musym:muname))


## Export
write_csv(backExtract_30, here('data/swd/a_polycarpa/backExtract_a_polycarpa_30cm_lowFilter.csv'))
write_csv(backExtract_100, here('data/swd/a_polycarpa/backExtract_a_polycarpa_100cm_lowFilter.csv'))
write_csv(backExtract_200, here('data/swd/a_polycarpa/backExtract_a_polycarpa_200cm_lowFilter.csv'))
write_csv(backExtract_full, here('data/swd/a_polycarpa/backExtract_a_polycarpa_full_lowFilter.csv'))
```



## Correlation

Assessing the correlation between environmental variables for SDM. Additional derived variables, such as mean temp and difference of temp, are generated. Then a Pearson correlation matrix is produced; based on the values, certain variables may be removed from the final SWD file.

*NOTE:* CHANGE THIS to be only for background pts 
```{r}
# testOcc <- read_csv(here('data/swd/occExtract_2000_2022.csv')) %>% 
#   mutate(tmean = (tmx+tmn)/2,
#          tdiff = tmx-tmn) %>% 
#   dplyr::select(aet:tdiff) %>% 
#   dplyr::select(!c(cwd, pet))
#
# occCor = cor(testOcc, method = "pearson", use = "complete.obs")
# corrplot(occCor, method = 'number', type = 'upper')


testBack <- read_csv(here('data/swd/a_polycarpa/backExtract_a_polycarpa_2000_2022_lowFilter5km.csv')) %>% 
  dplyr::select(aet:cec) %>% 
  mutate(tmean = (tmx+tmn)/2,
         tdiff = tmx-tmn) %>%
  dplyr::select(!c(cwd, pet, tmn, tmn))


cor <- cor(testBack, method = "pearson", use = "everything")
corrplot(cor, method = 'number', type = 'upper')

# summer <- testBack %>% 
#   dplyr::select(!ppt_winter) %>% 
#   filter(!is.na(tmx_summer))
# 
# summerCor = cor(summer, method = "pearson", use = "everything")
# corrplot(summerCor, method = 'number', type = 'upper')
# 
# winter <- testBack %>% 
#   dplyr::select(!tmx_summer) %>% 
#   filter(!is.na(ppt_winter))
# 
# winterCor = cor(winter, method = "pearson", use = "everything")
# corrplot(winterCor, method = 'number', type = 'upper')
```




# Maxent Models

Construct model using Samples with Data (SWD) method.


### Input data

Read in species occurrence and background data, then wrangle to put in proper format. The model will be run using two temperature predictors: mean temperature [(tmn + tmx)/2] and difference of temperature [tmx-tmn]. After these values are derived, both occurrence and background points will isolated by month and merged into one data frame per month .
```{r}
## Occurrence data ----------------------------------------------------------
swdOcc <- read_csv(here('data/swd/a_polycarpa/a_polycarpa_200cm_lowFilter.csv')) %>% 
  ## Add tmean, tdiff, and presence values
  mutate(tdiff = round(tmx-tmn, 2),
         presence = 1) %>% 
  ## remove select variables
  dplyr::select(!c(id, source, tmn, pet, cwd, ppt, date, ppt_winter_mean)) %>% 
  rename(x = lon, y = lat)

## rearrange x and y to match background df
swdOcc <- swdOcc[, c(2, 1, 4, 3, 5:13)]


## Background Data ----------------------------------------------------------
swdBack <- read_csv(here('data/swd/a_polycarpa/backExtract_a_polycarpa_200cm_lowFilter.csv')) %>% 
  mutate(tdiff = round(tmx-tmn, 2),
         presence = 0) %>% 
  dplyr::select(!c(tmn, pet, ppt, cwd, ppt_winter_mean))


## Bind into one df and
## remove any rows in NAs
swd <- rbind(swdOcc, swdBack)
swd <- swd[complete.cases(swd), ]

write_csv(swd, here('data/swd/a_polycarpa/swd_a_polycarpa_200cm_lowFilter.csv'))
```

### Model fitting
```{r}
## Read in all points
swd <- read_csv(here('data/swd/swd_2000_2022.csv'))

## Select env predictors for model (presence + background)
x <- swd %>% 
  dplyr::select(aet:tdiff)

## Specify occurrence data
p <- swd %>% 
  dplyr::select(presence)

## Arguments/options to pass to Maxent()
args <- c('jackknife=TRUE', 
          'autofeature=TRUE', 
          'responsecurves=TRUE', 
          'linear=TRUE',
          'quadratic=TRUE',
          'threshold=FALSE',
          'hinge=FALSE',
          'replicates=5', 'replicatetype=crossvalidate')

## Path to save results
path <- here('data/maxent_outputs/fitting/')



## Run test model
testModel <- maxent(x, p, path=path, args=args, removeDuplicates = TRUE)
```


### Final model
```{r}
## Read in all points
swd <- read_csv(here('data/swd/a_polycarpa/swd_a_polycarpa_full_lowFilter.csv'))

## Select env predictors for model (presence + background)
x <- swd %>% 
  dplyr::select(aet:tdiff)

## Specify occurrence data
p <- swd %>% 
  dplyr::select(presence)

## Set Arguments/Options to Pass to Maxent
args <- c('jackknife=TRUE', 
          'autofeature=TRUE', 
          'responsecurves=TRUE', 
          'linear=TRUE',
          'quadratic=TRUE',
          'product=TRUE',
          'threshold=FALSE',
          'hinge=FALSE', 
          'maximumiterations=100000', 
          'writeplotdata=TRUE')

## Path to save results
path <- here('data/maxent_outputs/a_polycarpa/lowFilter/full_depth')

#Final Model Creation
testFinal <- maxent(x, p, path=path, args=args)


# save(testFinal, file = here("data/maxent_outputs/final/Model.rData"))
```



## Predictions

Now, we want to project the Maxent model predictions for each month to determine how suitability changes during the year. 


### Average monthly environmental data:

First we need to create rasters representing the monthly average for each predictor variable used in the model.
```{r}
## Read in fxn
source(here('R/env_avg.R'))

## Run fxn (shows progress bar)
env_avg(var_names = c('aet', 'ppt', 'tmn', 'tmx'),
        pathIn = here('data/bcmv8/2000_2022//'),
        pathOut = here('data/bcmv8/monthly_avgs//'))
```


### Run by month:

Testing out prediction maps using only one year of monthly env data. 
NOTE; testing out how to make this fxn work with user defined name/amount of variables w/o having to manually change the function... will think more on later.
```{r}
variables = c('aet', 'ppt', 'tmn', 'tmx')
df <- as.data.frame(variables) %>% 
  mutate(varNum = seq(1:length(variables))) %>% 
  pivot_wider(names_from = varNum, names_prefix = "var", values_from = variables) %>% 
  slice(rep(1:n(), each = 12)) %>% 
  mutate(month = c('jan', 'feb', 'mar', 'apr', 'may', 'jun',
                   'jul', 'aug', 'sep', 'oct', 'nov', 'dec'))






## Read in fxn
source(here('R/pred_month.R'))

## Read in Maxent model
load(here('data/maxent_outputs/final/Model.rData'))

## Run fxn (shows progress bar)
pred_month(model = testFinal,
           pathIn = here('data/bcmv8/monthly_avgs//'),
           pathOut = here('data/maxent_outputs/predictions//'))
```



## Response Curves

```{r}

```


























