---
title: "Kern Natives SDM"
author: "Nick McManus"
date: "2023-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)    ## always
library(here)         ## reading/writing data
library(terra)        ## Better/faster GIS
library(raster)       ## GIS format required by Dismo
library(dismo)        ## Maxent pkg
library(lubridate)    ## Dates and progress bar
```


## Overview

This script extracts environmental data for each species occurrence (based on location and time). This data is then used to perform a species distribution model using Maxent with `dismo`.

*   Species occurrence data was pulled from GBIF (and maybe CalFlora?) and filtered to return reasonably accurate presence-only data in California. The script for this is in "src" directory.

*   Environmental data comes from the Basin Characterization Model version 8.0 (Flint et al., 2021) hosted by USGS. 



## Convert ASC to TIF

Publicly available BCMv8 files are in .asc format. ASCII files have associated coordinates but usually do not have an explicitly defined CRS when read in as a raster. Additionally, .tif have smaller file sizes than .asc, so conversion saves on data storage. Use this code to convert an entire file of .asc files to .tif files with appropriate CRS. BCM data uses NAD83 CA Albers (EPSG:3310)
```{r}
## Assign file path (selects all .asc files in directory)
file_asc <- list.files(path = here("data/bcmv8/2000_2009/"), 
                   pattern='\\.asc$', 
                   full=TRUE)
## Saves files w/same name but as .tif
file_tif <- gsub("\\.asc$", ".tif", file_asc)


## Progress bar
fileLength <- length(file_asc)
pb <- txtProgressBar(min = 0,
                     max = fileLength,
                     style = 3,
                     width = fileLength,
                     char = "=")


## Loop to read in .asc, add crs, then output .tif
for (i in 1:fileLength) {
    r <- rast(file_asc[i])
    crs(r) <- "epsg: 3310"
    writeRaster(r, file_tif[i])
    
    setTxtProgressBar(pb, i)
}
```



## Extract environmental data


Using BCMv8 rasters, we'll extract information for each species occurrence.
```{r}
## Read in occ df
spOcc <- read_csv(here('data/species_occurrence_GBIF/a_polycarpa.csv')) %>% 
  janitor::clean_names() %>% 
  dplyr::select(gbifid, decimallatitude, decimallongitude, month, year)


## Create df for dates
dates_df <- data.frame(year = rep(2021:2022, each = 12), 
                       mon = c('jan', 'feb', 'mar', 'apr', 'may', 'jun',
                               'jul', 'aug', 'sep', 'oct', 'nov', 'dec'),
                       mon_num = seq(1, 12, 1))
## Rearrange to water year format
## put last 3 mo at top, adjust calendar year
dates_df <- rbind(tail(dates_df, 3), head(dates_df, -3))
dates_df[1:3, 1] = (dates_df[4,1] - 1)
```

```{r}
## empty df to store loop results
extract_df <- data.frame()

## Extract loop
for (i in 1:nrow(dates_df)) {
  ## Read in list of filenames
  files <- list.files(path = here("data/bcmv8/2021_2022/"), 
                      ## only list those with matching yr/mo in name
                      pattern= paste0(dates_df[i, 1], dates_df[i, 2]), 
                      full=TRUE)
  
  ## Stack all rasters
  env_stack <- terra::rast(c(files))
  
  ## Filter obs to yr/mo
  spOcc_filter <- spOcc %>% 
    filter(year == dates_df[i,1],
           month == dates_df[i,3])

  ## If filtered df has obs, then vectorize and extract.
  if(nrow(spOcc_filter)>0) {
      ## vectorize and reproj to env data crs
      spOcc_vect <- spOcc_filter %>%
        vect(geom = c("decimallongitude", "decimallatitude"), crs = "WGS84") %>%
        project(y = crs(env_stack))

      ## extract and tidy df
      spExtract <- extract(env_stack, spOcc_vect, method = "simple") %>%
        ## only keep first 3 chars of each column name
        ## (e.g. "cwd2021jan" becomes "cwd")
        rename_with(~substr(., 1, 3)) %>%
        ## merge occ data w/extract data
        cbind(spOcc_filter, .) %>%
        dplyr::select(-ID)

      ## append results to df
      extract_df <- rbind(extract_df, spExtract)
    
    } else {
    ## If no obs for a yr/mo, skip
    next
  }

}
```





## Maxent



